The evolution of gene expression in the term placenta of viviparous mammals

### 2016-01-20 manuscript/manuscript.Rnw
discuss core placenta transcriptome
+ make 'the annexin complex' annexin complexes, since there's many
possible annexin complexes.

For the term placenta of \smonkey, assembled contigs from RNA-seq were
aligned using against \human and the best alignment was chosen as the
annotation for each assembled contig. In the term placenta of
\smonkey, the top 50 non-housekeeping genes
(\cref{tab:top50_spider_monkey_placenta}) include a possible regulator
of androgen receptor signaling
\cite{Sampson.ea2012:PAGE4positivityisassociated}, \textit{PAGE4} %
(\Sexpr{gene.expression("ateles_fusciceps","PAGE4")}), %
the pregnancy specific glycoprotein \textit{PSG8}
(\Sexpr{gene.expression("ateles_fusciceps","PSG8")}), %
a member of the chorionic somatotropins, \textit{GH1} %
(\Sexpr{gene.expression("ateles_fusciceps","GH1")}), %
a few HLA genes (\textit{HLA-A}, \textit{HLA-F},
\cref{tab:top50_spider_monkey_placenta}), and multiple components of
the annexin complex
(\Sexpr{gene.expression("ateles_fusciceps","ANXA2",complete=TRUE)}, %
\Sexpr{gene.expression("ateles_fusciceps","S100P",complete=TRUE)}, %
\Sexpr{gene.expression("ateles_fusciceps","S100A11",complete=TRUE)}) %
which may play a role in angiogenesis, and certainly regulate
cell-cell interaction in the developing placenta
\cite{Liu.ea2015:AnnexinA2complexeswith}.
\Sexpr{gene.expression("ateles_fusciceps","AL365202.1",complete=TRUE)}
is also highly expressed, has no known function, and is duplicated
multiple times in \textit{Macaca mulatta}. All genes with expression
greater than 50 \ac{FPKM} are listed in Supplementary
\cref{tab:all_ateles_fusciceps_placenta}.

@@

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \galili are shown in
\cref{tab:top50_n_galili_placenta} with a list of all highly expressed
genes in Supplemental \cref{tab:all_nannospalax_galili_placenta}.
Components of the annexin complex are highly expressed
(\Sexpr{gene.expression("nannospalax_galili","S100a6",complete=TRUE)},
\Sexpr{gene.expression("nannospalax_galili","S100a11",complete=TRUE)},
\Sexpr{gene.expression("nannospalax_galili","S100a10",complete=TRUE)},
and
\Sexpr{gene.expression("nannospalax_galili","Anxa2",complete=TRUE)}),
as are members of the thymosin family
(\Sexpr{gene.expression("nannospalax_galili","Tmsb4x",complete=TRUE)}
and
\Sexpr{gene.expression("nannospalax_galili","Tmsb10",complete=TRUE)}),
various cytoskeleton related proteins
(\Sexpr{gene.expression("nannospalax_galili","Actc1",complete=TRUE)},
\Sexpr{gene.expression("nannospalax_galili","Actg2",complete=TRUE)},
\Sexpr{gene.expression("nannospalax_galili","Acta2",complete=TRUE)},
\Sexpr{gene.expression("nannospalax_galili","Myl9",complete=TRUE)},
\Sexpr{gene.expression("nannospalax_galili","Vim",complete=TRUE)}), and
transmembrane protein of currently unknown function, 
\Sexpr{gene.expression("nannospalax_galili","Gm10134",complete=TRUE)}.

@@

The top 50 expressed genes as measured by \ac{FPKM} in the term
placenta of the \cow are shown in \cref{tab:top50_cow_placenta} with a
list of all genes in Supplemental \cref{tab:all_bos_taurus_placenta}.
The gene with the highest expression was \textit{TMSB10} %
(\Sexpr{gene.expression("bos taurus","TMSB10")}), and
\textit{TMSB4} was also highly expressed %
(\Sexpr{gene.expression("bos taurus","TMSB4")}). As in
\dog{}, members of the annexin complex were also highly expressed
(\Sexpr{gene.expression("bos taurus","S100G",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","ANXA2",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","S100A2",complete=TRUE)}, and %
\Sexpr{gene.expression("bos taurus","S100A11",complete=TRUE)}), as well as %
at least one of a set of four miRNA with 100\% identity
(\Sexpr{gene.expression("bos taurus","ENSBTAG00000047046",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","ENSBTAG00000047154",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","ENSBTAG00000045710",complete=TRUE)}, and %
\Sexpr{gene.expression("bos taurus","ENSBTAG00000046351",complete=TRUE)}). 
%
As these are novel miRNA, there is a possibility that the assignment
of the short reads to these four miRNA is due to miss-annotation of
the \cow genome.
% 
Several members of the ruminant-specific pregnancy-associated
glycoprotein family show especially high expression in \cow placenta
(\cref{tab:top50_cow_placenta}, \cref{fig:pag2_evolution_tree}, and
Supplemental \cref{tab:all_bos_taurus_placenta}) including
%
\Sexpr{gene.expression("bos taurus","PAG2",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","PAG10",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","ENSBTAG00000022759",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","ENSBTAG00000013214",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","ENSBTAG00000019226",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","ENSBTAG00000024611",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","PAG12",complete=TRUE)}. %
%
28 members of this family have $FPKM \ge 1$, and this family of
genes has greatly expanded in ruminants (\cow and \sheep{}) although
the expression levels of this family are higher in \cow than in \sheep
\cref{fig:pag2_evolution_tree}. Other genes with high expression
include genes involved with regulation of growth (\textit{IGF2},
\textit{PHLDA2}), imprinted genes (\textit{DLK1}, \textit{PHLDA2}),
and \textit{CSH1}, a gene that is a member of a ruminant specific
expansion of placental prolactins and not directly related to the gene
of the same name in hominids \ref{fig:prolactin_evolution_tree}. Other
ruminant-specific placental prolactins also show high expression in
the cow placenta including %
\Sexpr{gene.expression("bos taurus","PRP1",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","PRP5",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","PRP6",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","PRP8",complete=TRUE)}, and %
\Sexpr{gene.expression("bos taurus","PRP7",complete=TRUE)}, %
with a total of 14 with $\mathrm{FPKM} \ge 1$.
%
\textit{SSLP1} (\textit{SOLD1} in \sheep{}), a gene expressed in the
cotyledonary villi of the ruminiant placenta and involved in cell
proliferation
\cite{Ushizawa.ea2009:Characterizationandexpressionanalysis}, also
shows evidence of high expression %
(\Sexpr{gene.expression("bos taurus","SSLP1")}). %
There is some placental expression of the estrogen synthase gene
(\textit{CYP19A1}) in cows %
(\Sexpr{gene.expression("bos taurus","CYP19A1")}), and
comparatively greater expression of \textit{HSD3B} %
(\Sexpr{gene.expression("bos taurus","HSD3B")}).

@@

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \textit{Sus scrofa} are shown in
\cref{tab:top50_pig_placenta} with a list of all highly expressed
genes in Supplemental \cref{tab:all_sus_scrofa_placenta}. The top two
expressed transcripts,
\Sexpr{gene.expression("sus_scrofa","ENSSSCG00000025162",complete=TRUE)}
and
\Sexpr{gene.expression("sus_scrofa","ENSSSCG00000021356",complete=TRUE)},
are miRNA which correspond to the \cow miRNA
\href{http://www.mirbase.org/cgi-bin/mirna_entry.pl?acc=MI0013080}{bta-mir-2904-1}
and
\href{http://www.mirbase.org/cgi-bin/mirna_entry.pl?acc=MIMAT0013845}{bta-mir-2887-1},
respectively, both of which are also expressed in \cow ovary
\cite{Hossain.ea2009:Identificationandcharacterizationof}, although
neither of which have an obvious target gene on the basis of
\textit{in silico} analysis
\cite{Agarwal.ea2015:PredictingeffectivemicroRNAtarget}. Other highly
expressed genes include members of the insulin growth factor family
(\Sexpr{gene.expression("sus_scrofa","IGF2",complete=TRUE)} and
\Sexpr{gene.expression("sus_scrofa","IGFBP3",complete=TRUE)}), members
of the annexin complex
(\Sexpr{gene.expression("sus_scrofa","S100A1",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","ANXA2",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","S100A12",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","S100A10",complete=TRUE)}, and
\Sexpr{gene.expression("sus_scrofa","S100A1",complete=TRUE)}), the
galectin \Sexpr{gene.expression("sus_scrofa","LGALS3",complete=TRUE)}
and multiple cytoskeleton and extracellular matrix proteins
(\Sexpr{gene.expression("sus_scrofa","KRT19",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","COL1A1",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","BETA-ACTIN",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","KRT9",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","CLDN4",complete=TRUE)}, and
\Sexpr{gene.expression("sus_scrofa","KRT9",complete=TRUE)}).

@@

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \textit{Equus Caballus} are shown in
\cref{tab:top50_horse_placenta} with a list of all highly expressed
genes in Supplemental \cref{tab:all_equus_caballus_placenta}. The gene
with the highest expression was the relaxin,
\Sexpr{gene.expression("equus_caballus","RLN",complete=TRUE)},
followed by two miRNA
(\Sexpr{gene.expression("equus_caballus","ENSECAG00000025328",complete=TRUE)}
and
\Sexpr{gene.expression("equus_caballus","ENSECAG00000027056",complete=TRUE)})
with no known \textit{in silico} targets
\cite{Agarwal.ea2015:PredictingeffectivemicroRNAtarget}, the thymosin
\Sexpr{gene.expression("equus_caballus","TMSB4",complete=TRUE)},
members of the insulin growth factor family
(\Sexpr{gene.expression("equus_caballus","IGF2",complete=TRUE)} and
\Sexpr{gene.expression("equus_caballus","IGFBP3",complete=TRUE)}) and
multiple members of the annexin complex
(\Sexpr{gene.expression("equus_caballus","S100A6",complete=TRUE)},
\Sexpr{gene.expression("equus_caballus","S100A1",complete=TRUE)}, and
\Sexpr{gene.expression("equus_caballus","ANXA2",complete=TRUE)}),
similar to the pattern of expression seen in \pig and \dog.

@@

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \sheep are shown in
\cref{tab:top50_sheep_placenta} with a list of all highly expressed
genes in Supplemental \cref{tab:all_ovis_aries_placenta}. The complete
list of highly expressed genes includes many snRNAs and rRNAs likely
due to the RNA extraction and library preparation procedure utilized
for this sample. Ignoring these, notable highly expressed genes
include members of the annexin complex
(\Sexpr{gene.expression("ovis_aries","S100G",complete=TRUE)},
\Sexpr{gene.expression("ovis_aries","ANXA2",complete=TRUE)}),
hemoglobin subunits
(\Sexpr{gene.expression("ovis_aries","ENSOARG00000010944",complete=TRUE)}),
an ortholog of \textit{SSLP1},
\Sexpr{gene.expression("ovis_aries","SORD1",complete=TRUE)}, two
cathespins (\Sexpr{gene.expression("ovis_aries","CTSL",complete=TRUE)}
and \Sexpr{gene.expression("ovis_aries","CTSB",complete=TRUE)}), and
placentally expressed transcript 1
\Sexpr{gene.expression("ovis_aries","PLET1",complete=TRUE)} which may
play a role in cellular migration
\cite{Raymond.ea2010:Expressionoforphanprotein}.

@@

The
\Sexpr{core.placenta.transcriptome.long[,length(unique(human_name))]}
genes comprising the core placenta transcriptome are shown in
\cref{fig:core_placenta_transcriptome} ordered by their median
transcription level in \ac{FPKM}. We considered a gene to be core to
the placenta transcriptome if a 1:1 ortholog of that gene was
expressed with $\mathrm{FPKM} \ge 10$ in all species studied and was
not a human housekeeping gene according to
\citet{Eisenberg.ea2013:Humanhousekeepinggenesrevisited}.
\Cref{fig:core_placenta_transcriptome_shape} shows the number of genes
which would be included in the core placenta set if the \ac{FPKM}
cutoff was set differently.
%
Multiple components of the annexin complex are included in the core
transcriptome, including \textit{ANXA2}, \textit{ANXA1},
\textit{S100A11}, \textit{S100A10}. The modulation of  \DLA{It does more stuff; read
  the huthcinson2011 review and fill this in}
%
% Another annexin, \textit{ANXA2} 
%
\textit{TIMP3}, which may be down-regulated in pre-eclampsia
\cite{Ma.ea2014:Down-regulationofTIMP3leads} (though another study
\cite{Xiang.ea2013:PromoterhypomethylationofTIMP3} claims
up-regulation) does things

@@

SERPINH1? with TIMP3 
% http://www.proteinatlas.org/ENSG00000149257-SERPINH1/tissue/placenta

\Cref{tab:core_placenta_go_terms} lists the significant enrichment of
\ac{GO} terms in the core placenta transcriptome; notable terms
include those involved in cytoskeletal structure, protein transport,
and membrane protein structure.

>>>>

For the term placenta of \smonkey, assembled contigs from RNA-seq were
aligned using against \human and the best alignment was chosen as the
annotation for each assembled contig. In the term placenta of
\smonkey, the top 50 non-housekeeping genes
(\cref{tab:top50_spider_monkey_placenta}) include a possible regulator
of androgen receptor signaling
\cite{Sampson.ea2012:PAGE4positivityisassociated}, \textit{PAGE4} %
(\Sexpr{gene.expression("ateles_fusciceps","PAGE4")}), %
the pregnancy specific glycoprotein \textit{PSG8}
(\Sexpr{gene.expression("ateles_fusciceps","PSG8")}), %
a member of the chorionic somatotropins, \textit{GH1} %
(\Sexpr{gene.expression("ateles_fusciceps","GH1")}), %
a few HLA genes (\textit{HLA-A}, \textit{HLA-F},
\cref{tab:top50_spider_monkey_placenta}), and multiple components of
annexin complexes
(\Sexpr{gene.expression("ateles_fusciceps","ANXA2",complete=TRUE)}, %
\Sexpr{gene.expression("ateles_fusciceps","S100P",complete=TRUE)}, %
\Sexpr{gene.expression("ateles_fusciceps","S100A11",complete=TRUE)}) %
which may play a role in angiogenesis, and certainly regulate
cell-cell interaction in the developing placenta
\cite{Liu.ea2015:AnnexinA2complexeswith}.
\Sexpr{gene.expression("ateles_fusciceps","AL365202.1",complete=TRUE)}
is also highly expressed, has no known function, and is duplicated
multiple times in \textit{Macaca mulatta}. All genes with expression
greater than 50 \ac{FPKM} are listed in Supplementary
\cref{tab:all_ateles_fusciceps_placenta}.

@@

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \galili are shown in
\cref{tab:top50_n_galili_placenta} with a list of all highly expressed
genes in Supplemental \cref{tab:all_nannospalax_galili_placenta}.
Components of annexin complexes are highly expressed
(\Sexpr{gene.expression("nannospalax_galili","S100a6",complete=TRUE)},
\Sexpr{gene.expression("nannospalax_galili","S100a11",complete=TRUE)},
\Sexpr{gene.expression("nannospalax_galili","S100a10",complete=TRUE)},
and
\Sexpr{gene.expression("nannospalax_galili","Anxa2",complete=TRUE)}),
as are members of the thymosin family
(\Sexpr{gene.expression("nannospalax_galili","Tmsb4x",complete=TRUE)}
and
\Sexpr{gene.expression("nannospalax_galili","Tmsb10",complete=TRUE)}),
various cytoskeleton related proteins
(\Sexpr{gene.expression("nannospalax_galili","Actc1",complete=TRUE)},
\Sexpr{gene.expression("nannospalax_galili","Actg2",complete=TRUE)},
\Sexpr{gene.expression("nannospalax_galili","Acta2",complete=TRUE)},
\Sexpr{gene.expression("nannospalax_galili","Myl9",complete=TRUE)},
\Sexpr{gene.expression("nannospalax_galili","Vim",complete=TRUE)}), and
transmembrane protein of currently unknown function, 
\Sexpr{gene.expression("nannospalax_galili","Gm10134",complete=TRUE)}.

@@

The top 50 expressed genes as measured by \ac{FPKM} in the term
placenta of the \cow are shown in \cref{tab:top50_cow_placenta} with a
list of all genes in Supplemental \cref{tab:all_bos_taurus_placenta}.
The gene with the highest expression was \textit{TMSB10} %
(\Sexpr{gene.expression("bos taurus","TMSB10")}), and
\textit{TMSB4} was also highly expressed %
(\Sexpr{gene.expression("bos taurus","TMSB4")}). As in
\dog{}, members of annexin complexes were also highly expressed
(\Sexpr{gene.expression("bos taurus","S100G",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","ANXA2",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","S100A2",complete=TRUE)}, and %
\Sexpr{gene.expression("bos taurus","S100A11",complete=TRUE)}), as well as %
at least one of a set of four miRNA with 100\% identity
(\Sexpr{gene.expression("bos taurus","ENSBTAG00000047046",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","ENSBTAG00000047154",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","ENSBTAG00000045710",complete=TRUE)}, and %
\Sexpr{gene.expression("bos taurus","ENSBTAG00000046351",complete=TRUE)}). 
%
As these are novel miRNA, there is a possibility that the assignment
of the short reads to these four miRNA is due to miss-annotation of
the \cow genome.
% 
Several members of the ruminant-specific pregnancy-associated
glycoprotein family show especially high expression in \cow placenta
(\cref{tab:top50_cow_placenta}, \cref{fig:pag2_evolution_tree}, and
Supplemental \cref{tab:all_bos_taurus_placenta}) including
%
\Sexpr{gene.expression("bos taurus","PAG2",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","PAG10",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","ENSBTAG00000022759",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","ENSBTAG00000013214",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","ENSBTAG00000019226",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","ENSBTAG00000024611",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","PAG12",complete=TRUE)}. %
%
28 members of this family have $FPKM \ge 1$, and this family of
genes has greatly expanded in ruminants (\cow and \sheep{}) although
the expression levels of this family are higher in \cow than in \sheep
\cref{fig:pag2_evolution_tree}. Other genes with high expression
include genes involved with regulation of growth (\textit{IGF2},
\textit{PHLDA2}), imprinted genes (\textit{DLK1}, \textit{PHLDA2}),
and \textit{CSH1}, a gene that is a member of a ruminant specific
expansion of placental prolactins and not directly related to the gene
of the same name in hominids \ref{fig:prolactin_evolution_tree}. Other
ruminant-specific placental prolactins also show high expression in
the cow placenta including %
\Sexpr{gene.expression("bos taurus","PRP1",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","PRP5",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","PRP6",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","PRP8",complete=TRUE)}, and %
\Sexpr{gene.expression("bos taurus","PRP7",complete=TRUE)}, %
with a total of 14 with $\mathrm{FPKM} \ge 1$.
%
\textit{SSLP1} (\textit{SOLD1} in \sheep{}), a gene expressed in the
cotyledonary villi of the ruminiant placenta and involved in cell
proliferation
\cite{Ushizawa.ea2009:Characterizationandexpressionanalysis}, also
shows evidence of high expression %
(\Sexpr{gene.expression("bos taurus","SSLP1")}). %
There is some placental expression of the estrogen synthase gene
(\textit{CYP19A1}) in cows %
(\Sexpr{gene.expression("bos taurus","CYP19A1")}), and
comparatively greater expression of \textit{HSD3B} %
(\Sexpr{gene.expression("bos taurus","HSD3B")}).

@@

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \textit{Sus scrofa} are shown in
\cref{tab:top50_pig_placenta} with a list of all highly expressed
genes in Supplemental \cref{tab:all_sus_scrofa_placenta}. The top two
expressed transcripts,
\Sexpr{gene.expression("sus_scrofa","ENSSSCG00000025162",complete=TRUE)}
and
\Sexpr{gene.expression("sus_scrofa","ENSSSCG00000021356",complete=TRUE)},
are miRNA which correspond to the \cow miRNA
\href{http://www.mirbase.org/cgi-bin/mirna_entry.pl?acc=MI0013080}{bta-mir-2904-1}
and
\href{http://www.mirbase.org/cgi-bin/mirna_entry.pl?acc=MIMAT0013845}{bta-mir-2887-1},
respectively, both of which are also expressed in \cow ovary
\cite{Hossain.ea2009:Identificationandcharacterizationof}, although
neither of which have an obvious target gene on the basis of
\textit{in silico} analysis
\cite{Agarwal.ea2015:PredictingeffectivemicroRNAtarget}. Other highly
expressed genes include members of the insulin growth factor family
(\Sexpr{gene.expression("sus_scrofa","IGF2",complete=TRUE)} and
\Sexpr{gene.expression("sus_scrofa","IGFBP3",complete=TRUE)}), members
of annexin complexes
(\Sexpr{gene.expression("sus_scrofa","S100A1",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","ANXA2",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","S100A12",complete=TRUE)}, and
\Sexpr{gene.expression("sus_scrofa","S100A10",complete=TRUE)}), the
galectin \Sexpr{gene.expression("sus_scrofa","LGALS3",complete=TRUE)}
and multiple cytoskeleton and extracellular matrix proteins
(\Sexpr{gene.expression("sus_scrofa","KRT19",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","COL1A1",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","BETA-ACTIN",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","KRT9",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","CLDN4",complete=TRUE)}, and
\Sexpr{gene.expression("sus_scrofa","KRT9",complete=TRUE)}).

@@

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \textit{Equus Caballus} are shown in
\cref{tab:top50_horse_placenta} with a list of all highly expressed
genes in Supplemental \cref{tab:all_equus_caballus_placenta}. The gene
with the highest expression was the relaxin,
\Sexpr{gene.expression("equus_caballus","RLN",complete=TRUE)},
followed by two miRNA
(\Sexpr{gene.expression("equus_caballus","ENSECAG00000025328",complete=TRUE)}
and
\Sexpr{gene.expression("equus_caballus","ENSECAG00000027056",complete=TRUE)})
with no known \textit{in silico} targets
\cite{Agarwal.ea2015:PredictingeffectivemicroRNAtarget}, the thymosin
\Sexpr{gene.expression("equus_caballus","TMSB4",complete=TRUE)},
members of the insulin growth factor family
(\Sexpr{gene.expression("equus_caballus","IGF2",complete=TRUE)} and
\Sexpr{gene.expression("equus_caballus","IGFBP3",complete=TRUE)}) and
multiple members of annexin complexes
(\Sexpr{gene.expression("equus_caballus","S100A6",complete=TRUE)},
\Sexpr{gene.expression("equus_caballus","S100A1",complete=TRUE)}, and
\Sexpr{gene.expression("equus_caballus","ANXA2",complete=TRUE)}),
similar to the pattern of expression seen in \pig and \dog.

@@

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \sheep are shown in
\cref{tab:top50_sheep_placenta} with a list of all highly expressed
genes in Supplemental \cref{tab:all_ovis_aries_placenta}. The complete
list of highly expressed genes includes many snRNAs and rRNAs likely
due to the RNA extraction and library preparation procedure utilized
for this sample. Ignoring these, notable highly expressed genes
include members of annexin complexes
(\Sexpr{gene.expression("ovis_aries","S100G",complete=TRUE)},
\Sexpr{gene.expression("ovis_aries","ANXA2",complete=TRUE)}),
hemoglobin subunits
(\Sexpr{gene.expression("ovis_aries","ENSOARG00000010944",complete=TRUE)}),
an ortholog of \textit{SSLP1},
\Sexpr{gene.expression("ovis_aries","SORD1",complete=TRUE)}, two
cathespins (\Sexpr{gene.expression("ovis_aries","CTSL",complete=TRUE)}
and \Sexpr{gene.expression("ovis_aries","CTSB",complete=TRUE)}), and
placentally expressed transcript 1
\Sexpr{gene.expression("ovis_aries","PLET1",complete=TRUE)} which may
play a role in cellular migration
\cite{Raymond.ea2010:Expressionoforphanprotein}.

@@

The
\Sexpr{core.placenta.transcriptome.long[,length(unique(human_name))]}
genes comprising the core placenta transcriptome are shown in
\cref{fig:core_placenta_transcriptome} ordered by their median
transcription level in \ac{FPKM}. We considered a gene to be core to
the placenta transcriptome if a 1:1 ortholog of that gene was
expressed with $\mathrm{FPKM} \ge 10$ in all species studied and was
not a human housekeeping gene according to
\citet{Eisenberg.ea2013:Humanhousekeepinggenesrevisited}.
\Cref{fig:core_placenta_transcriptome_shape} shows the number of genes
which would be included in the core placenta set if the \ac{FPKM}
cutoff was set differently.
%
Multiple components of annexin complexes are included in the core
transcriptome, including \textit{ANXA2}, \textit{ANXA1},
\textit{S100A11}, \textit{S100A10}. ANXA1 is involved in the
resolution of the inflammation response through inhibition of
phospholipase A2 activity and signaling through the formyl peptide
receptor family, and has previously been hypothesized to be important
for the maintenance of the anti-inflammatory during pregnancy
\cite{Hutchinson.ea2011:Molecularregulatorsofresolution}. ANXA2 is
known to interact with S100A10, S100A6, S100A11, and S100A4
\cite{Liu.ea2015:AnnexinA2complexeswith}, and there is evidence for
its interaction with S100P as well; it is involved in cell-cell
interaction as well as vesicle trafficking and von Willibrand Factor
secretion. The expression regulation of another core gene,
\textit{TIMP3}, has been implicated in pre-eclampsia
\cite{Ma.ea2014:Down-regulationofTIMP3leads,Xiang.ea2013:PromoterhypomethylationofTIMP3}.
\Cref{tab:core_placenta_go_terms} lists the significant enrichment of
\ac{GO} terms in the core placenta transcriptome; notable terms
include those involved in cytoskeletal structure, protein transport,
membrane protein structure, and inflammation resolution.

@@

% SERPINH1? with TIMP3 
% http://www.proteinatlas.org/ENSG00000149257-SERPINH1/tissue/placenta
